---
title: "lab12"
output: html_document
date: "2022-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidycensus)
library(tidyverse)
```

# Finding variable
## Microdata

```{r}
person_vars2018 <- pums_variables %>%
  distinct(year, survey, var_code, var_label, data_type, level) %>%
  filter(level=="person", year == 2018, survey=="acs5")

head(person_vars2018)
```


```{r}
hh_vars2018 <- pums_variables %>%
  distinct(year, survey, var_code, var_label, data_type, level) %>%
  filter(level=="housing", year == 2018, survey=="acs1")

head(hh_vars2018)
```


## Area level data

```{r}
area_vars_2018 <- load_variables(2018, "acs1")
head(area_vars_2018)
```


```{r}
area_vars_2010 <- load_variables(2010, "sf1")
head(area_vars_2010)
```

# Loading variables
```{r, echo=FALSE}
person_data <- get_pums(variables=c("TEN", "SEX", "PINCP", "WKHP","JWMNP"), state = "TX", year = 2018, survey="acs1", recode=TRUE) %>%
  filter(JWMNP>0)
```


```{r}
person_data <- person_data %>% subset(select=c(TEN_label, SEX_label, PINCP, WKHP, JWMNP))

person_data
```



```{r}
knitr::kable(person_data[1:10,])
```
```{r, echo=FALSE}
hh_data <- get_pums(variables=c("TEN", "YBL", "BDSP", "GRNTP","BLD","LNGI"), state = "NY", year = 2018, survey="acs1", recode=TRUE) %>%
  filter(TEN_label=="Rented") %>%
  mutate(age_struct= case_when(
    YBL_label=="1939 or earlier" ~ 100,
    YBL_label=="1940 to 1949" ~ 73,
    YBL_label=="1950 to 1959" ~ 63,
    YBL_label=="1960 to 1969" ~ 53,
    YBL_label=="1970 to 1979" ~ 43,
    YBL_label=="1970 to 1989" ~ 33,
    YBL_label=="1990 to 1999" ~ 23,
    YBL_label=="2000 to 2004" ~ 16,
    YBL_label=="2005" ~ 13,
    YBL_label=="2006" ~ 12,
    YBL_label=="2007" ~ 11,
    YBL_label=="2008" ~ 10,
    YBL_label=="2009" ~ 9,
    YBL_label=="2010" ~ 8,
    YBL_label=="2011" ~ 7,
    YBL_label=="2012" ~ 6,
    YBL_label=="2013" ~ 5,
    YBL_label=="2014" ~ 5,
    YBL_label=="2015" ~ 3,
    YBL_label=="2016" ~ 2,
    YBL_label=="2017" ~ 1,
    YBL_label=="2018" ~ 0,)) %>%
  subset(select=c(age_struct, BDSP, GRNTP, BLD_label, LNGI_label))
```


```{r}
hh_data
```



```{r}
pop_vars <- c(workers_16 = "B23025_002",
          unemp_pop_16 = "B23025_005",
          college_25 = "B23006_023",
          popln_25 = "B23006_001",
          med_income = "B06011_001",
          tot_pop = "B01003_001",
          white_pop = "B01001A_001")

rural_vars <- c(total_hh = "H002001",
                rural_hh = "H002005")

counties2018 <- get_acs(geography = "county", variables=pop_vars, output="wide", year=2018, survey="acs1")

counties_rural <- get_decennial(geography = "county", variables = rural_vars, output = "wide", year=2010)

counties <- inner_join(counties2018, counties_rural, by="GEOID") %>%
  mutate(pct_college=college_25E/popln_25E,
         unemp_rate=unemp_pop_16E/workers_16E,
         majority_white=white_popE, tot_popE > 0.5,
         rural=rural_hh==total_hh) %>%
  subset(select=c(pct_college, unemp_rate, majority_white, rural, med_incomeE))
  
counties

```
# My turn 1

```{r}
tracts_2019 <- get_acs(geography ="tract", variables = pop_vars, state = "CA", county = "Los Angeles", output = "wide", year = 2019, survey = "acs5"  )

tracts <- tracts_2019 %>%
  mutate(pct_college = college_25E / popln_25E,
         unemp_rate = unemp_pop_16E / workers_16E,
         majority_white = white_popE / tot_popE > 0.5) %>%
  subset(select=c(GEOID, pct_college, unemp_rate, majority_white, med_incomeE))


tracts

```



# Downloading TIGER Shapefiles


```{r, echo=FALSE}
library(tigris)
library(sf)

urban <- urban_areas()
```


```{r}
cbus <- urban %>%
  st_transform(26917) %>%
  filter(str_detect(NAMELSAD10, "Columbus, OH"))
```


```{r}
```

# My turn 2

```{r}
LA <- urban %>%
  st_transform(26917) %>%
  filter(UACE10== 51445)

LA_sf <- st_as_sf(LA)

library(tmap)
tmap_mode("view")
tm_shape(LA_sf) + tm_polygons("UACE10")

```

# My turn 3

I have found that the urban areas and metro divisions boundaries are different for Boston area. So, I used both boundaries to show the water bodies. Since Boston is speaded over five counties, I filtered water bodies of those five counties and intersected that with the two areal boundaries mentioned above. In the below code chunk, I first showed with the metro division.  

```{r, echo=FALSE}
#Loading boston metropolitan area
place <- metro_divisions()
```


```{r, echo=FALSE}
#filtering metro division that has boston, MA in their names
boston <- place %>%
  st_transform(26917) %>%
  filter(str_detect(NAMELSAD, "Boston, MA"))


# loading and joining water bodies of three counties in MA that is within Boston

waterN <- area_water(state = "MA", county = "Norfolk")
waterS <- area_water(state = "MA", county = "Suffolk")
waterP <- area_water(state = "MA", county = "Plymouth")
waterE <- area_water(state = "MA", county = "Essex")
waterM <- area_water(state = "MA", county = "Middlesex")

water <- rbind(waterN,waterP,waterS, waterE,waterM) %>% st_transform(26917)

water1 <- water[boston, ]

```


```{r}
#plot both


tm_shape(boston) + tm_borders("black", lwd = 2.5)+
  tm_shape(water1) + tm_fill("orange")
```


The following chunk shows output for Boston urban areas 
```{r}

boston2 <- urban %>%
  st_transform(26917) %>%
  filter(str_detect(NAMELSAD10, "Boston, MA"))

water2 <- water[boston2, ]

tm_shape(boston2) + tm_borders("black", lwd = 1.5)+
  tm_shape(water2) + tm_fill("orange")
```


# Downloading data from OSM

```{r}
library(dplyr)
library(tmap)
library(sp)
library(stringi)
library(osmextract)

dir.create("Boston data")
SLO_file <- oe_match("Boston")
SLO_streets <- oe_read(SLO_file$url, provider = "openstreetmap_fr", download_directory = "Boston data", layer= "lines", quiet = TRUE) %>%
  filter(!is.na(highway))

```

# My Turn 4

```{r}
#first take the boundary of Cleveland urban area
cleveland <- urban %>%
  st_transform(26917) %>%
  filter(str_detect(NAMELSAD10, "Cleveland, OH"))

tm_shape(cleveland) + tm_borders()
# Now download the streets of Cleveland
dir.create("Cleveland data")
SLO_fileC <- oe_match("Cleveland")
SLO_streetsC <- oe_read(SLO_fileC$url, provider = "openstreetmap_fr", download_directory = "Cleveland data", layer= "lines", quiet = TRUE) %>%
  filter(!is.na(highway)) %>% 
  st_transform(26917)

# Clip the streets that fall under the Cleveland boundary

SLO_streetsC <- SLO_streetsC[cleveland, ]

```

# My Turn 5

```{r}

dir.create("Cleveland data")
SLO_fileC <- oe_match("Cleveland")

SLO_pointsC <- oe_read(SLO_fileC$url, provider = "openstreetmap_fr", download_directory = "Cleveland data", layer= "lines", quiet = TRUE) %>%
  filter(str_detect(other_tags, '"amenity"=>"bicycle_parking"'))%>%
  st_transform(26917)


head(SLO_pointsC$other_tags)
#mapping parking data

tm_shape(SLO_pointsC) + tm_lines("red",lwd = 5)
```

         
         
